Stochastic Approximation Monte Carlo with Dynamic Update Factor (SAD)

### 2018-08-13 papers/histogram/sad.tex
describe sad from last friday.

Outside of the energy range $E_L \le E \le E_H$, the weights are ignored
when computing acceptance probabilities, and the weights $w(E_L)$ or
$w(E_H)$ are used instead.  This means that the histogram is \emph{not}
flat in this region, but instead is proportional to the density of
states.  In order to ensure that $w(E) \propto D(E)$, the change in
weight with each visit should be equal to the change in weight at
the border energy $E=E_{H/L}$.
\begin{align}
  w_{t+1}(E)-w_t(E) &= w_{t+1}(E_{H/L}) -w_t(E_{H/L}) \\
    &= (e^{\gamma_t}-1) w_t(E_{H/L})
\end{align}
This ensures that although we visit each energy above $E_H$ or
below $E_L$ only
in proportion to its density of states, the weights at these energies
will approach the entropy as the simulation progresses.

@@

Energy binning is an important consideration for complex continous
systems. Coarse binning yields fewer energy states for the walker to
explore but more time spent in each individual bin. Fine binning yields
more energy states to explore with less time spent in each bin.
\davidsays{move this discussion earlier.} Some
methods such as AdaWL~\cite{koh2013dynamically} (discussed elsewhere)
employ a tunable mechanism for controlling the binning for low entropic
states in order to ensure the exploration of all energies. A
significant advantage of SAD over other flat histogram methods is that
energy binning is solved dynamically. SAD should perform independently
for a reasonable range of binning because $\gamma \propto
\nicefrac{N_{E}}{t}$.  As the number of energy states found $N_{E}$
increases (fine binning), the time spent $t$ in each bin will decrease.
This argument explains why SAD should perform well for complex systems
in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
similar way but this adds yet another parameter for the user to choose.

>>>>

Outside of the energy range $E_L \le E \le E_H$, the weights are
determined from the accumulated energy histogram. For each energy $E_i
> E_H$, the energy weights are calculated
\begin{align}
w(E_i)_{t+1} = w(E_H)_{t} + \ln\left(\frac{H(E_i)}{H_{\max}}\right)
\end{align}
assuming of course that the simulation has collected statistics at that energy.
For each energy $E_i
< E_L$, the energy weights are calculated
\begin{align}
w(E_i)_{t+1} = w(E_L)_{t} + \ln\left(\frac{H(E_i)}{H_{\max}}\right) +
\frac{E_i - E_L}{T_{\min}}
\end{align}

@@

A significant advantage of SAD over other flat histogram methods is
that energy binning is solved dynamically. SAD should perform
independently for a reasonable range of binning because $\gamma \propto
\nicefrac{N_{E}}{t}$.  As the number of energy states found $N_{E}$
increases (fine binning), the time spent $t$ in each bin will decrease.
This argument explains why SAD should perform well for complex systems
in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
similar way but this adds yet another parameter for the user to choose.
